---
title: "p8105_hw2_xw2598"
author: "Xinyao Wu"
date: "2018-10-02"
output: github_document
---
##problem1

```{r}
library(tidyverse)
nyc_data = read.csv("./data files/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")%>% 
  #data clean
janitor::clean_names() %>% 
  #retain specific variables 
select(line:entry,vending,ada) %>% 
  #convert the variable entry from character to logic
 #gather(key="routes_served", value = "route_value",route1:route11 ) %>% 
mutate(
     entry=as.character(entry),
     entry = replace(entry,entry =="YES", "TRUE"),
       entry =replace(entry,entry =="NO","FALSE"),
       entry = as.logical(entry)
)
```

```{r}
#new dataset variables
names(nyc_data) 
#The dimension of the resulting dataset
print("rows:")
nrow(nyc_data)
print("columns:")
ncol(nyc_data)

```

This dataset contains variables:
line,station name,station latitude,station longitude,route1,route2,route3,route4,route5,route6,route7,route8,route9,route10,route11,entrance type,entry,vending,ada. My data cleaning steps: 1.Change the variables names. 2.Select the specific variables according to the requirement. 3.Change the type of the data in "entry". The dimension of the resulting dataset is 1868 rows * 19 columns.These data is not tidy enough, I suggest to gather all of the routes into one variable so I put related codes in the comment lines.

```{r}
##The number of distinct stations:
#select the variables that can identify stations
n_station = select(nyc_data,line,station_name,ada) %>% 
distinct()
nrow(n_station)
##The number of stations that are ADA compliant
ada_compliant = filter(n_station, ada=="TRUE") 
  nrow(ada_compliant)
##The proportion of station entrances / exits without vending allow entrance
nrow(filter(nyc_data,vending=="NO",entry=="TRUE"))/nrow(filter(nyc_data,vending=="NO"))

```
There are 465 distinct stations, and there are 84 stations that are ADA compliant.The proportion of station entrances / exits without vending allow entrance is 0.377.

```{r}
#Reformat data so that route number and route name are distinct variables
new_nyc_data = nyc_data %>% 
gather(key="route_name", value = "route_number",route1:route11 ) %>% 
#Number of Dinstinct stations that serve the A train
select(line,station_name, route_name, route_number,ada) %>% 
  #filte A train
  filter(route_number == "A") 
  #count distinct station serving A train
  A_station = distinct(new_nyc_data,line,station_name,.keep_all = TRUE) 
  nrow(A_station)
# carculate number  of  ADA compliant in stations serving A train
  nrow(filter(A_station,ada=="TRUE"))
# carculate proportion of  ADA compliant in stations serving A train
  nrow(filter(A_station,ada=="TRUE"))/nrow(A_station)
```
60 distinct stations that serve the A train
Of the stations that serve the A train, the number of ada compliant stations serving A train is , and the proportion of ADA compliant is 0.283.

##problem 2


Read and clean the Mr. Trash Wheel sheet:
```{r}
library(readxl)
#read data and omit columns containing notes
trash = read_excel("./data files/HealthyHarborWaterWheelTotals2017-9-26.xlsx",range = "A2:N258") %>% 
  #data clean 
  janitor::clean_names() %>% 
  #omit rows that do not include dumpster-specific data
  filter(dumpster!="NA")


```

@specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
@use reasonable variable names
@omit rows that do not include dumpster-specific data

rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)
Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2017? What was the median number of sports balls in a dumpster in 2016



